---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png",
  dpi = 300
)
```

# ffm

<!-- badges: start -->
<!-- badges: end -->

`{ffm}` is an R package that provides quick and easy access to data from
the geodata center of Germany's Federal Agency for Cartography and Geodesy (BKG).
The BKG is the official provider of spatial data in Germany and provides quite
a few datasets as open data. These data range from administrative areas to
earth observation data and are often crucial when working with regional
statistics from Germany. Part of the motivation for this package stems from
frustration when working areal identifiers returned by packages like
`{wiesbaden}` or `{restatis}` and not being able to quickly link them to their
spatial representations.

The name `ffm` is based on the colloquial short name of Frankfurt am Main where
the BKG's headquarters are located.

## Installation

You can install the development version of ffm from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jslth/ffm")
```

## Example

Retrieving data is pretty straightforward:

```{r simple_call}
library(ffm)
districts <- bkg_admin(level = "krs", scale = "5000")
```

<details>
<summary>
Code for the plot
</summary>
```{r, eval=FALSE}
library(ggplot2)

ggplot(districts) +
  geom_sf(fill = NA) +
  theme_void()
```
</details>

```{r, echo=FALSE}
library(ggplot2)

ggplot(districts) +
  geom_sf(fill = NA) +
  theme_void()
```

The package makes it easy to go deeper than just getting the data. In many
functions, you can use spatial filters.

```{r spat_filters}
districts <- bkg_admin(
  level = "krs",
  scale = "5000",
  bbox = c(xmin = 700000, ymin = 5900000, xmax = 850000, ymax = 6000000),
  predicate = "intersects"
)
districts$label <- ifelse(
  districts$bez == "Kreisfreie Stadt",
  paste("Stadt", districts$gen),
  districts$gen
)
```

<details>
<summary>
Code for the plot
</summary>
```{r, eval=FALSE}
library(ggrepel)

ggplot(districts) +
  geom_sf(fill = NA) +
  geom_text_repel(
    aes(label = label, geometry = geometry),
    stat = "sf_coordinates",
    size = 3
  ) +
  theme_void()
```
</details>

```{r, echo=FALSE}
library(ggrepel)

ggplot(districts) +
  geom_sf(fill = NA) +
  geom_text_repel(
    aes(label = label, geometry = geometry),
    stat = "sf_coordinates",
    size = 3
  ) +
  theme_void()
```

Attribute filters are supported using a R-like syntax.

```{r att_filters}
munics <- bkg_admin(
  level = "gem",
  cutoff = "3112",
  sn_l %in% c("08", "09")
)
```

<details>
<summary>
Code for the plot
</summary>
```{r, eval=FALSE}
munics$popdens <- munics$ewz / munics$kfl
munics$popdens[munics$popdens == 0] <- NA
ggplot(munics) +
  geom_sf(aes(fill = popdens), color = NA) +
  scale_fill_viridis_b("Population / km²", transform = "log10") +
  ggtitle("Population density in Southern Germany") +
  theme_void()
```
</details>

```{r, echo=FALSE}
library(ggplot2)

munics$popdens <- munics$ewz / munics$kfl
munics$popdens[munics$popdens == 0] <- NA
ggplot(munics) +
  geom_sf(aes(fill = popdens), color = NA) +
  scale_fill_viridis_b("Population / km²", transform = "log10") +
  ggtitle("Population density in Southern Germany") +
  theme_void()
```
